Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2014-03-23 Paper/gammaRays.tex
Removed model results from model description

%\begin{table}
%	\centering
%	\small
%	\begin{tabular}{ l | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3]}
%		\multicolumn{1}{ c |}{GRB Name} & $1\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
%		%\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
%		\hline
%		080916C	&	1.042	&	1.200	&	1.300	&	1.400	&	1.500
%	\end{tabular}
%	\caption{Burst data used in our study. The first row contains the corresponding \texttt{GRBurst} class variable names. The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}. \url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
%	\label{tab:observationResults}
%\end{table}

@@

The burst is fully specified by the following set of parameters:
\begin{itemize}
\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.

	If $\gamma$ is small enough, high photon densities would make the jet plasma optically thick to pair production. This in turn would suppress emitted flux, and this is not observed. From this argument the lower bound of $\gamma \ge 100$ was derived \cite{}. The limited total energy of the burst provides an upper bound: $\gamma \le \gamma_\text{max}$ (see section 3.6). Finally, the model dependent computations of $\gamma$ give the values ranging from $50$ for low-energy bursts to $1000$ for high-energy ones \cite{Ghirlanda:2011ux,Ghisellini:2012nc}. $\gamma$ may vary from burst to burst, and it should decrease during the burst due to the dissipation of energy. As a first approximation we select {\boldmath$\gamma = 100$} for this study.
}
\item{$\eta_0$, which defines the luminosity scale.

	$\eta_0$ controls the number of photons we see from the bursts, and so can be roughly estimated from the number of luminous bursts we observe. We can use this and the measured burst redshifts to estimate both $\eta_0$ and the bursts density $\rho_0$, which will be discussed later. For this study we choose {\boldmath$\eta_0 = 1 \times 10^{45}$}.
}
\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
\item{$\omega_0$, a characteristic radiation frequency;}
\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
\end{itemize}

@@

Our set of parameters led to a sample you can see on the figure \ref{fig:kappaDistribution}. However, the error margins of the observed stretching factors are too large to use the KS-test on them. Nevertheless, it appears that our model doesn't contradict existing observations.

\subsection{Very high energy bursts fraction}

Out final test is the comparison of answers to this question: given the bursts which were observed in a high energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a very high energy range $\left(\omega_2,\omega_3\right)$?

@@

The ratio for our set of parameters is $f_m = 0.072$. The LAT burst catalog \cite{Ackermann:2013zfa} contains $n = 35$ bursts, $3$ of which were observed in the energy range $\left(\unit[1]{GeV}, \infty\right)$. So the ratio is $f_o = \frac{3}{35} \approx 0.086$. The binary distribution gives us the probability for it to happen:
\begin{equation}
\text{Pr}\left(f_o = \frac{k}{n}\right) = \binom{n}{k} f_m^k \left(1-f_m\right)^{n-k} \approx 0.22
\end{equation}
which means that the model doesn't contradict observations in this test.

@@

>>>>

@@

The burst is fully specified by the following set of parameters:
\begin{itemize}
\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.}
\item{$\eta_0$, which defines the luminosity scale.}
\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
\item{$\omega_0$, a characteristic radiation frequency;}
\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
\end{itemize}

@@

With this algorithm, we arrive to our second test / constraint for the burst parameters. The obtained stretching factors distribution should be compatible with observations.

\subsection{High energy bursts fraction}

Out final test is the comparison of answers to this question: given the bursts which were observed in a whole energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a high energy range $\left(\omega_2,\omega_3\right)$?

@@

This is our 3rd test -- the ratio obtained should be compatible with the observed value.

@@

Finally, we are in a position to use the tests to fit parameters of the model:

\begin{itemize}
	\item{
		$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.

	}
	\item{$\eta_0$, which defines the luminosity scale.}
	\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
	\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
	\item{$\omega_0$, a characteristic radiation frequency;}
	\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
	\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
	\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
\end{itemize}
